---
title: "Dog Breeds"
date: "2 2 2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggimage)
```

## Context

<https://eu.usatoday.com/picture-gallery/life/2021/06/28/the-50-most-popular-dog-breeds-in-america/45134329/>

<https://www.vox.com/2016/8/31/12715176/most-popular-dog-breeds>

280 breeds

these dogs are registered through the [American Kennel Club](http://www.akc.org/dog-breeds/) (AKC), which categorizes and tallies up the totals of more than 150 breeds.

> In 1991, the Labrador retriever became the most popular dog --- and for the past 25 years it has retained its throne
>
> psychology professor Hal Herzog laid out what he called the "cultural
> drift" effect: Instead of studying up on various breeds, **dog owners tend
> to mimic the choices of others** --- neighbors, friends, celebrities.
>
> "**breeds enjoy heydays of approximately 25 years** ... [which] usually allows
> for two to three generations of dogs, as the breed transforms from
> novel to passé."
>
>
## R Import

```{r import}

breed_traits <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-02-01/breed_traits.csv')

trait_description <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-02-01/trait_description.csv')

breed_rank_all <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-02-01/breed_rank.csv')
```

## EDA

```{r}


Top <- breed_rank_all %>% 
  janitor::clean_names() %>% 
  select(-links) %>% 
  pivot_longer(x2013_rank:x2020_rank,
               names_to = "year", 
               values_to = "Rank") %>% 
  mutate(year = parse_number(year)) %>% 
  filter(year == 2020) %>% 
  slice_head(n = 10) 


# make breed_traits long
variables_long <- breed_traits %>% 
  select(-c(`Coat Type`, `Coat Length`)) %>% 
  pivot_longer(-Breed, 
               names_to = "variable", 
               values_to = "Trait_score") %>% 
  rename(breed = Breed) %>% 
  drop_na()

# Filter for Top 10 dogs in Year 2020

filtered <- variables_long %>% 
  full_join(Top, by = "breed") %>% 
  drop_na(image, Trait_score, variable) 
```

## Graphic

```{r}

 ggplot(filtered, 
       aes(x = variable, 
           y = breed, 
           fill = Trait_score)) +
  geom_tile(color = "white", 
            size = 0.2) +
  labs(
    title = "Trait Scores of Selected Dog breeds", 
    subtitle = "in the Year 2020"
  ) +
  scale_fill_viridis_b(
    direction = -1,
    n.breaks = 5, 
    option = "D", 
    name = "Trait Score",
    guide = guide_colorbar(
      frame.colour = "black",
      ticks.colour = "black",
      ticks.linewidth = grid::unit(1, "inch"),
      direction = "horizontal", 
      title.position = "top" 
      #label.position = "bottom"
    )
  )  + 
  scale_y_discrete(name = NULL,  # instead of axis.title = element_blank()
                   expand = c(0,0)) +
 # https://www.tidyverse.org/blog/2020/03/ggplot2-3-3-0/
  scale_x_discrete(name = NULL, 
                   expand = c(0,0), 
                   guide = guide_axis(n.dodge = 3)) +
  theme(
    plot.title = element_text(size = 14, 
                              face = "bold"),
    plot.subtitle = element_text(size = 12, 
                                 margin = margin(b = 5)),
    legend.position = "top", 
    axis.text = element_text(size = 08, 
                             color = "#222222"),
    axis.ticks = element_blank(), 
    plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"))
```

## Save 

```{r}

ggsave("DogsBreed.png")
```
